An Empirical Analysis of the Role of Energy in Economic Growth

### 2015-02-06 Articles/Article_EnerEcon_2014/EconGrowth4.Rnw
Adjusted the discussion of the CES boundary models.

****

In addition to the above trials, 
and in a manner similar to our approach to the Cobb-Douglas model,
we fit along all possible CES boundaries.
Boundaries are given by all combinations of the following:
$\delta_1$~=~0~or~1, 
$\sigma_1$~=~0~or~$\infty$,
$\delta$~=~0~or~1, and
$\sigma$~=~0~or~$\infty$.
Equation~\ref{eq:ces_generic} shows the generic three-factor CES equation,
wherein factors of production ($x_1$,~$x_2$,~$x_3$) 
are permutations of ($k$,~$l$,~$e$).
%
\begin{equation} \label{eq:ces_generic}
  y = \gamma A \left\{ \delta \left[ \delta_1 x_1^{-\rho_1} 
      + \left( 1 - \delta_1 \right) x_2^{-\rho_1} \right]^{\rho/\rho_1}  
          + \left( 1 - \delta \right) x_3^{-\rho}  \right\} ^ {-1/\rho}
\end{equation}
%
Table~\ref{tab:CES_boundary_models} shows the set 
of 20 degenerate equations found along parameter boundaries.
Each degenerate equation provides a boundary model.
Note that many boundary parameter combinations 
yield equivalent CES boundary models,%
  \footnote{
  For example, all possible boundary parameter combinations 
  that include $\delta = 0$ yield the same boundary model, $y = \gamma A x_3$.
  }
so there are far fewer than the maximum possible number of degenerate equations~(81).%
  \footnote{
  There are three possible states (unspecified, lower boundary, upper boundary)
  for four parameters ($\delta_1$, $\sigma_1$, $\delta$, $\sigma$),
  which gives $3^4 = 81$ possible boundary models.
  }
Boundary models obtained when $\delta_1$ or $\delta$ is at an extreme value
are straightforward to derive from Equation~\ref{eq:ces_generic}.
When $\sigma_1$ or $\sigma$ is 0, 
factors of production are complimentary,
and the Leontief model applies.%
  \footnote{
  E.g., when $\sigma_1 = 0$ and $\delta~=~1$, the Leontief model is
  $y = \gamma A \min (x_1, x_2)$.
  }
When $\sigma_1$ or $\sigma$ is $\infty$, 
factors of production are subsitutible,
and the linear model applies.%
  \footnote{
  E.g., when $\sigma_1 = \infty$ and $\delta~=~1$, the linear model is
  $y = \gamma A \left[ \delta_1 x_1 + (1 - \delta_1) x_2 \right]$.
  }
In the parameter columns of Table~\ref{tab:CES_boundary_models}, 
``---'' indicates that a parameter is unknowable on that boundary
(because it does not appear in the boundary model), and
an empty space indicates that the parameter 
can be estimated on that boundary using the degenerate equation.
The top portion of Table~\ref{tab:CES_boundary_models} shows 
two-parameter CES boundary models
(similar to the CES model without energy in Equation~\ref{eq:CES})
with $\delta~=~1$, $\sigma$~unknowable, and $x_3$~absent.
If the fitted parameters in a boundary model violate constraints, 
the boundary model is rejected.

\begin{table} \caption{Boundary models for the CES equation.} 
\label{tab:CES_boundary_models} 
  \begin{center}
    \begin{tabular}{r c c c c l} 
      \toprule
      & $\delta_1$  & $\sigma_1$ & $\delta$   & $\sigma$ & Boundary model \\
      \midrule
      1. & 1           & ---        & 1          & ---      & $y = \gamma A x_1$ \\
      2. & 0           & ---        & 1          & ---      & $y = \gamma A x_2$ \\
      3. & ---         & 0          & 1          & ---      & $y = \gamma A \min (x_1, \, x_2)$ \\
%
      4. &             & $\infty$   & 1          & ---        
            & $y = \gamma A \left[ \delta_1 x_1 + (1 - \delta_1) x_2 \right]$ \\
%
      5. &             &            & 1          & ---
            & $y = \gamma A \left[ \delta_1 x_1^{-\rho_1} 
                    + \left( 1 - \delta_1 \right) x_2^{-\rho_1} \right]^{-1/\rho_1}$ \\
\midrule
      6. & ---         & ---        & 0          & ---      & $y = \gamma A x_3$ \\
      7. & 1           & ---        & ---        & 0        & $y = \gamma A \min (x_1, \, x_3)$ \\
      8. & 0           & ---        & ---        & 0        & $y = \gamma A \min (x_2, \, x_3)$ \\
      9. & ---         & 0          & ---        & 0        & $y = \gamma A \min (x_1, \, x_2, \, x_3)$ \\
%
     10. & 1           & ---        &            & $\infty$        
            & $y = \gamma A \left[ \delta x_1 + (1 - \delta) x_3 \right]$ \\
%
     11. & 0           & ---        &            & $\infty$        
            & $y = \gamma A \left[ \delta x_2 + (1 - \delta) x_3 \right]$ \\
%
     12. &             & $\infty$   &            & $\infty$        
            & $y = \gamma A \left\{ \delta \left[ \delta_1 x_1 + \left( 1 - \delta_1 \right) x_2 \right] 
                    + \left( 1 - \delta \right) x_3  \right\}$ \\
%
     13. & ---         & 0          &            & $\infty$        
            & $y = \gamma A \left[ \delta \min (x_1, \, x_2) + \left( 1 - \delta \right) x_3  \right]$ \\
%
     14. &             & $\infty$   & ---        & 0
            & $y = \gamma A \min \left[ \delta_1 x_1 + \left(1 - \delta_1 \right) x_2, \, x_3  \right]$ \\
%
     15. & 1           & ---        &         & 
            & $y = \gamma A \left[ \delta x_1^{-\rho} 
                    + \left(1 - \delta \right) x_3^{-\rho} \right] ^ {-1/\rho}$ \\ 
%
     16. & 0           & ---        &            & 
            & $y = \gamma A \left[ \delta x_2^{-\rho} + \left(1 - \delta \right) x_3^{-\rho} \right]^{-1/\rho}$ \\
%
     17. &             & $\infty$   &            &
            & $y = \gamma A \left\{ \delta \left[ \delta_1 x_1 
                    + \left( 1 - \delta_1 \right) x_2 \right]^{-\rho}  
                    + \left( 1 - \delta \right) x_3^{-\rho}  \right\} ^ {-1/\rho}$ \\
%
     18. & ---         & 0          &            & 
            & $y = \gamma A \left\{ \delta \left[ \min \left( x_1, \, x_2 \right) \right]^{-\rho}  
                    + \left( 1 - \delta \right) x_3^{-\rho}  \right\} ^ {-1/\rho}$ \\ 
%
     19. &             &            &            & $\infty$
            & $y = \gamma A \left\{ \delta \left[ \delta_1 x_1^{-\rho_1} 
                    + \left( 1 - \delta_1 \right) x_2^{-\rho_1} \right]^{-1/\rho_1}  
                    + \left( 1 - \delta \right) x_3  \right\}$ \\ 
%
     20. &            &            & ---     & 0
            & $y = \gamma A \min \left\{ \left[ \delta_1 x_1^{-\rho_1} 
                    + \left( 1 - \delta_1 \right) x_2^{-\rho_1} \right]^{-1/\rho_1}, \, x_3 \right\}$ \\ 
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

****

The fit with lowest $sse$ of all above trials 
is deemed the winning model and presented in graphs.

\citet{Henningsen:2011td}, 
in their detailed analysis of \citet{Kemfert:1998vd}, found that

\begin{quote}
\ldots the Levenberg-Marquardt and the \texttt{PORT} algorithms 
are---at least in this study---most likely to find the 
coefficients that give the best fit to the model, 
where the \texttt{PORT} algorithm can be used to restrict the estimates 
to the economically meaningful region.
\end{quote}

In our testing, we found that to be mostly true. 
\texttt{PORT} nearly always provided lower
$sse$ than \texttt{L-BFGS-B}, despite the fact that 
\texttt{L-BFGS-B} often reports convergence
and \texttt{PORT} does not for the same data
(i.e., for the same $y$, $k$, $l$, and $e$ time series).

>>>>

\citet{Henningsen:2011td}, 
in their detailed analysis of \citet{Kemfert:1998vd}, found that 
%
\begin{quote}
\ldots the Levenberg-Marquardt and the \texttt{PORT} algorithms 
are---at least in this study---most likely to find the 
coefficients that give the best fit to the model, 
where the \texttt{PORT} algorithm can be used to restrict the estimates 
to the economically meaningful region.
\end{quote}
%
In our testing, we found that to be mostly true. 
\texttt{PORT} nearly always provided lower
$sse$ than \texttt{L-BFGS-B}, despite the fact that 
\texttt{L-BFGS-B} often reports convergence
and \texttt{PORT} does not for the same data
(i.e., for the same $y$, $k$, $l$, and $e$ time series).

In addition to the above trials, 
and in a manner similar to our approach to the Cobb-Douglas model,
we fit along all possible CES boundaries.
Boundaries are given by all combinations of the following:
$\delta_1$~=~0~or~1, 
$\delta$~=~0~or~1,
$\sigma_1$~=~0~or~$\infty$, and
$\sigma$~=~0~or~$\infty$.
Table~\ref{tab:CES_boundary_models} shows the set 
of 20 degenerate equations found along parameter boundaries~(rows~1--20).
Row~21 of Table~\ref{tab:CES_boundary_models} shows the generic three-factor CES equation,
wherein factors of production ($x_1$,~$x_2$,~$x_3$) 
are permutations of ($k$,~$l$,~$e$).
Each degenerate equation provides a boundary model.
Note that many boundary parameter combinations 
yield equivalent CES boundary models,%
  \footnote{
  For example, all possible boundary parameter combinations 
  that include $\delta = 0$ yield the same boundary model, $y = \gamma A x_3$.
  }
so there are far fewer than the maximum possible number of degenerate equations~(81).%
  \footnote{
  There are three possible states (unspecified, lower boundary, upper boundary)
  for four parameters ($\delta_1$, $\sigma_1$, $\delta$, $\sigma$),
  which gives $3^4 = 81$ possible boundary models.
  }
Boundary models obtained when $\delta_1$ or $\delta$ is at an extreme value
are straightforward to derive from row~21 of Table~\ref{tab:CES_boundary_models}.
When $\sigma_1$ or $\sigma$ is 0, 
factors of production are perfect complements,
and the Leontief model applies.%
  \footnote{
  E.g., when $\sigma_1 = 0$ and $\delta~=~1$, the Leontief model is
  $y = \gamma A \min (x_1, x_2)$.
  }
When $\sigma_1$ or $\sigma$ is $\infty$, 
factors of production are perfect substitutes,
and the linear model applies.%
  \footnote{
  E.g., when $\sigma_1 = \infty$ and $\delta~=~1$, the linear model is
  $y = \gamma A \left[ \delta_1 x_1 + (1 - \delta_1) x_2 \right]$.
  }
In the parameter columns of Table~\ref{tab:CES_boundary_models}, 
``---'' indicates that a parameter is unknowable on that boundary
(because it does not appear in the boundary model), and
an empty space indicates that the parameter 
can be estimated on that boundary using the degenerate equation.
The top portion of Table~\ref{tab:CES_boundary_models} shows 
two-parameter CES boundary models
(similar to the CES model without energy in Equation~\ref{eq:CES})
with $\delta~=~1$, $\sigma$~unknowable, and $x_3$~absent.
If the fitted parameters in a boundary model violate constraints, 
the boundary model is rejected.

\begin{table} \caption{Boundary (rows~1--20) and generic (row~21) models for the CES equation.} 
\label{tab:CES_boundary_models} 
  \begin{center}
    \begin{tabular}{r c c c c l} 
      \toprule
      & $\delta_1$  & $\sigma_1$ & $\delta$   & $\sigma$ & Boundary model \\
      \midrule
      1. & 1           & ---        & 1          & ---      & $y = \gamma A x_1$ \\
      2. & 0           & ---        & 1          & ---      & $y = \gamma A x_2$ \\
      3. & ---         & 0          & 1          & ---      & $y = \gamma A \min (x_1, \, x_2)$ \\
%
      4. &             & $\infty$   & 1          & ---        
            & $y = \gamma A \left[ \delta_1 x_1 + (1 - \delta_1) x_2 \right]$ \\
%
      5. &             &            & 1          & ---
            & $y = \gamma A \left[ \delta_1 x_1^{-\rho_1} 
                    + \left( 1 - \delta_1 \right) x_2^{-\rho_1} \right]^{-1/\rho_1}$ \\
\midrule
      6. & ---         & ---        & 0          & ---      & $y = \gamma A x_3$ \\
      7. & 1           & ---        & ---        & 0        & $y = \gamma A \min (x_1, \, x_3)$ \\
      8. & 0           & ---        & ---        & 0        & $y = \gamma A \min (x_2, \, x_3)$ \\
      9. & ---         & 0          & ---        & 0        & $y = \gamma A \min (x_1, \, x_2, \, x_3)$ \\
%
     10. & 1           & ---        &            & $\infty$        
            & $y = \gamma A \left[ \delta x_1 + (1 - \delta) x_3 \right]$ \\
%
     11. & 0           & ---        &            & $\infty$        
            & $y = \gamma A \left[ \delta x_2 + (1 - \delta) x_3 \right]$ \\
%
     12. &             & $\infty$   &            & $\infty$        
            & $y = \gamma A \left\{ \delta \left[ \delta_1 x_1 + \left( 1 - \delta_1 \right) x_2 \right] 
                    + \left( 1 - \delta \right) x_3  \right\}$ \\
%
     13. & ---         & 0          &            & $\infty$        
            & $y = \gamma A \left[ \delta \min (x_1, \, x_2) + \left( 1 - \delta \right) x_3  \right]$ \\
%
     14. &             & $\infty$   & ---        & 0
            & $y = \gamma A \min \left[ \delta_1 x_1 + \left(1 - \delta_1 \right) x_2, \, x_3  \right]$ \\
%
     15. & 1           & ---        &         & 
            & $y = \gamma A \left[ \delta x_1^{-\rho} 
                    + \left(1 - \delta \right) x_3^{-\rho} \right] ^ {-1/\rho}$ \\ 
%
     16. & 0           & ---        &            & 
            & $y = \gamma A \left[ \delta x_2^{-\rho} + \left(1 - \delta \right) x_3^{-\rho} \right]^{-1/\rho}$ \\
%
     17. &             & $\infty$   &            &
            & $y = \gamma A \left\{ \delta \left[ \delta_1 x_1 
                    + \left( 1 - \delta_1 \right) x_2 \right]^{-\rho}  
                    + \left( 1 - \delta \right) x_3^{-\rho}  \right\} ^ {-1/\rho}$ \\
%
     18. & ---         & 0          &            & 
            & $y = \gamma A \left\{ \delta \left[ \min \left( x_1, \, x_2 \right) \right]^{-\rho}  
                    + \left( 1 - \delta \right) x_3^{-\rho}  \right\} ^ {-1/\rho}$ \\ 
%
     19. &             &            &            & $\infty$
            & $y = \gamma A \left\{ \delta \left[ \delta_1 x_1^{-\rho_1} 
                    + \left( 1 - \delta_1 \right) x_2^{-\rho_1} \right]^{-1/\rho_1}  
                    + \left( 1 - \delta \right) x_3  \right\}$ \\ 
%
     20. &            &            & ---     & 0
            & $y = \gamma A \min \left\{ \left[ \delta_1 x_1^{-\rho_1} 
                    + \left( 1 - \delta_1 \right) x_2^{-\rho_1} \right]^{-1/\rho_1}, \, x_3 \right\}$ \\ 
\midrule
     21. &            &            &         & 
            & $y = \gamma A \left\{ \delta \left[ \delta_1 x_1^{-\rho_1} 
                    + \left( 1 - \delta_1 \right) x_2^{-\rho_1} \right]^{\rho/\rho_1}  
                    + \left( 1 - \delta \right) x_3^{-\rho}  \right\} ^ {-1/\rho}$ \\ 
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

The model with lowest $sse$ of all above trials 
is deemed the winning (i.e.,~best) model,
and its parameters are presented in graphs herein.
In special cases, two boundary models will be equivalent.
For example, boundary model~3 is identical to boundary model~1~(2)
if $x_1$ is always less than (greater than) $x_2$ for all years.
If two or more models have the same $sse$ (to six digits), 
the model with the lowest row number in Table~\ref{tab:CES_boundary_models} is deemed the winner,
thereby ensuring that the simplest model is selected.

**** Edit the next paragraphs after we finalize the appearance
of the distribution graphs. 
Also, add a discussion of the boundary models and how the reduce 
the amount of data available in the bar graphs. ****
